---
title: "MultiStart parameter sensitivity analysis"
author: "Pratik A. Khare"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r, echo = FALSE}
load("C:/PRATIK/PhD/Mike Betenbaugh/Hybrid model project/Manuscript/Review/R1 MjC2.3/Multiple initial start sensitivity analysis/KOM and SOM sensitivity analysis outputs.RData")
```

## Fval (objective value) comparison between KOM and SOM

First, we will just see the distribution of the fval for the KOM and SOM side-by-side via a boxplot:

```{r, echo=FALSE}
plot1 <- boxplot(
     as.numeric(MultiStart_KOM_output$fval),as.numeric(MultiStart_SOM_output$fval),
     ylab="Fval",
     main="Spread of objective function values for KOM and SOM",
     names = c("KOM", "SOM"),
     col = c("darkgreen", "orange"),
     border = c("black", "black"))
```

As seen in the boxplot, the KOM and SOM have distinctly different fval values.  
&rarr; The range for KOM fvals is `r range(MultiStart_KOM_output$fval)`  
&rarr; The range for SOM fvals is `r range(MultiStart_SOM_output$fval)`  

The difference between the fvals for the KOM and SOM is significant, as seen by the p-value: `r t.test(MultiStart_KOM_output$fval, MultiStart_SOM_output$fval)$p.value`

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
